Elephant Seal Age JAE Intraspecific Variation

### 2024-05-01 paper/paperv3.qmd
Adding sections to check out recapture rate

resight <- read.csv(here::here("data/raw/fullresights.csv"))

@@

resight <- resight %>% 
  dplyr::select(animalID, resightID, season, date, timeofyear, calyear) %>% 
  mutate(animalID = as.factor(animalID))

full <- semi_join(resight, sealdat, "animalID") %>% group_by(animalID, calyear) %>% count(calyear)

ggplot(full, aes(calyear, n)) + 
  geom_point()

>>>>

resight <- read.csv(here::here("data/raw/fullresights.csv")) %>% 
  dplyr::select(animalID, resightID, season, date, timeofyear, calyear, yday) %>% 
  mutate(animalID = as.factor(animalID))

@@

full <- semi_join(resight, sealdat, "animalID") %>% 
  mutate(fiscalyear = ifelse(yday > 200, calyear + 1, calyear)) %>% 
  group_by(animalID, fiscalyear) %>% 
  count(fiscalyear) %>% 
  mutate(year = as.numeric(fiscalyear)) %>% 
  filter(year > 1987)

full %>% 
  group_by(year) %>% 
  summarize(across(n, list(median = median, 
                           q25 = \(x) quantile(x, 0.25), 
                           q75 = \(x) quantile(x, 0.75)))) %>% 
  ggplot(aes(year, n_median)) +
  geom_ribbon(aes(ymin = n_q25, ymax = n_q75), fill = "grey80") +
  geom_line() +
  geom_hline(yintercept = 3, linetype = "dashed", color = "firebrick") +
  labs(y = "Resights per year") + 
  theme_classic()

#some animals have multiple rows because they have multiple pups associated with them...
sealdat %>% 
  group_by(animalID, year) %>% 
  mutate(n = n()) %>% 
  filter(n > 2)

skips <- sealdat %>% 
  distinct(animalID, year) %>% 
  group_by(animalID) %>% 
  summarize(n_obs = n(),
            obs_window = max(year) - min(year) + 1,
            skips = obs_window - n_obs) %>% 
  group_by(obs_window, skips) %>% 
  summarize(ind = n(), .groups = "drop")

resights_per_year <- full %>% 
  group_by(year) %>% 
  summarize(median_resights = median(n))

skips_per_year <- skips

skip_frac <- sealdat %>% 
  distinct(animalID, year) %>% 
  mutate(year = as.numeric(as.character(year))) %>% 
  group_by(animalID) %>% 
  summarize(firstobs = min(year),
            lastobs = max(year)) %>% 
  mutate(n = lastobs - firstobs + 1) %>% 
  uncount(n) %>% 
  group_by(animalID) %>% 
  mutate(year = firstobs + seq(n()) - 1) %>% 
  ungroup() %>% 
  mutate(year = factor(year)) %>% 
  left_join(transmute(sealdat, animalID, year, skip = 0), by = c("animalID", "year")) %>% 
  mutate(skip = ifelse(is.na(skip), 1, 0)) %>% 
  group_by(year) %>% 
  summarize(skip_frac = sum(skip) / n())

resights_per_year %>% 
  mutate(year = factor(year)) %>% 
  left_join(skip_frac, by = "year") %>% 
  mutate(year = as.numeric(as.character(year))) %>% 
  ggplot(aes(median_resights, skip_frac)) +
  geom_point(aes(fill = year), shape = 21, color = "grey20", size = 2) +
  ggrepel::geom_label_repel(aes(label = year), max.overlaps = 8) +
  scale_y_continuous("Skip fraction", labels = scales::percent) +
  scale_fill_distiller(palette = "PuRd") +
  labs(x = "Median resights") +
  theme_classic()

#mark recapture model is necessary? could have bred at another colony. Many skips!!

ggplot(skips, aes(obs_window, skips, size = ind)) +
  geom_point()
